# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data
### load DF
```{r}
DF <- read.csv("activity.csv")
```

### clean date
```{r CLEANDATE}
DF$DateGood<-as.Date(DF$date)
```

## What is mean total number of steps taken per day?
### Sum steps per day
```{r Sumstepsperday}
DailySteps <- as.vector(by(DF$steps, DF$DateGood, function(x) sum(x, na.rm = TRUE)))
```

### make the histogram
```{r make the histogram}
hist(DailySteps, breaks = 20, xlab = "Daily steps", col = "red", main = "Freq of Daily Steps")
```

### mean + median
```{r mean + median}
MeanMedianSteps<-data.frame(mean = mean(DailySteps), median = median(DailySteps))
MeanMedianSteps
```


## What is the average daily activity pattern?

### Calc the ave steps + 5MIN intervals
```{r Calc the ave steps + 5MIN intervals }
AveStepsTaken <- as.vector(by(DF$steps, DF$interval, function(x) mean(x, na.rm = TRUE)))
MinuteIntervals <- levels(as.factor(DF$interval))
```

### make the plot
```{r ### make the plot}
plot(MinuteIntervals, AveStepsTaken, type = "l", col="red",xlab = "5 min Interval number", ylab = "Ave number daily steps", main = "5-minute interval + average number steps taken, averaged across all days")
```

### find interval with max steps
```{r find interval with max steps}
MaxInterval<-as.numeric(MinuteIntervals[which.max(AveStepsTaken)] ) ; MaxInterval
```
### Combine ave steps and 5minintervals
```{r Combine ave steps and 5minintervals}
DFX<-data.frame(MinuteIntervals,AveStepsTaken)
```
### NUmber steps in MaxInterval
```{r NUmber steps in MaxInterval}
DFX[(MinuteIntervals==MaxInterval),]
```

## Imputing missing values

### for intervals with NA I am going to replace NA with the average steps for that 5-minute interval
### Calc # rows with NA
```{r Calc # rows with NA}
NaRowsCount<-nrow(DF[is.na(DF$steps),]) ; NaRowsCount
```

### Create a new data set with missing data replaced, i call it DFClean
### Add the calculated ave steps for 5MinIntervals through left-join
```{r Add the calculated ave steps for 5MinIntervals through left-join}
DFClean<-merge(x = DF, y = DFX, by.x = "interval", by.y = "MinuteIntervals", all.x=TRUE)
```

### Now pick ave steps for steps = NA
```{r Now pick ave steps for steps is NA }
DFClean$CleanSteps<-ifelse(is.na(DFClean$steps)== T,DFClean$AveStepsTaken,DFClean$steps)
```

### make the cleaned hist

### Sum clean steps per day
```{r Sum clean steps per day}
CleanDailySteps <- as.vector(by(DFClean$CleanSteps, DFClean$DateGood, function(x) sum(x, na.rm = TRUE)))
```
### make the cleaned hist
```{r make the cleaned hist}
hist(CleanDailySteps, breaks = 20, xlab = "Daily steps", col = "red", main = "Freq of Cleaned Daily Steps")
```

### clean mean + median
```{r clean mean + median}
CleanMeanMedianSteps<-data.frame(CleanMean = mean(CleanDailySteps), Cleanmedian = median(CleanDailySteps))
CleanMeanMedianSteps
```

### Total number of steps has increased since we replaced NAs with the average for that interval.
### Appararantly this method created a mean and median that is exactly the same, wonder if this is because we put the averages in and so it trended towards that



## Are there differences in activity patterns between weekdays and weekends?

### Add weekday and then what type of day that is
```{r  Add weekday and then what type of day that is}
DFClean <- cbind(DFClean, DayType = factor(weekdays(DFClean$DateGood) == "Saturday" | weekdays(DFClean$DateGood) == "Sunday", labels = c("weekday", "weekend")))
```

### make the plot
```{r make the plot}
par(mfrow = c(2, 1))


means.weekday <- as.vector(by(DFClean[DFClean$DayType == "weekday",]$CleanSteps, DFClean[DFClean$DayType == "weekday",]$interval, mean))
means.weekend <- as.vector(by(DFClean[DFClean$DayType == "weekend",]$CleanSteps, DFClean[DFClean$DayType == "weekend",]$interval, mean))

plot(MinuteIntervals, means.weekday, type = "l",ylim=c(0,250), xlab = "Interval", ylab = "Mean daily steps", main = "Weekdays")
plot(MinuteIntervals, means.weekend, type = "l",ylim=c(0,250), xlab = "Interval", ylab = "Mean daily steps", main = "Weekends")
```





